Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down

### 2022-08-30 paper/ms.tex
Add sam section

\newcommand{\RCR}{\ensuremath{R_{\textrm{CR}}}}
\newcommand{\Rot}{\ensuremath{\mathcal{R}}}
\newcommand{\Vc}{\ensuremath{V_{\textrm{c}} }}
\newcommand{\PS}{\ensuremath{\Omega_{\textrm{p}}}}
\newcommand{\Rb}{\ensuremath{R_{\textrm{b}}}}

@@

\section{Methods}
\label{sec:methods}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{fig/ps_torque.pdf}
    \caption{The \textit{upper panel} shows the evolution of the pattern speed.
    As expected, the bar in the \Nbody{} run slows down due to interactions
    between the bar and the dark matter halo. However, the bar in the SMUGGLE
    run does not slow down and instead remains at a constant pattern speed. The
    \textit{lower panel} shows the torque on the bar by different components.
    The solid lines indicate the torque exerted by the halo in both the \Nbody{}
    and SMUGGLE cases. The dashed line is the torque exerted by the gas phase in
    the SMUGGLE run (there is no gas in the \Nbody{} run). Details on how these
    properties are calculated is given in the Methods section.}\label{fig:prop}
\end{figure}

@@

\subsection{Bar Analysis}
The analysis of various bar properties is performed as follows. First, the
pattern speed is measured from the angle of the second Fourier component. We
measured the second Fourier component by computing,
\begin{equation}
\begin{split}
A_2 &= \sum_i m_i e^{i 2 \phi_i} \\
A_0 &= \sum_i m_i \textrm{,}
\end{split}
\end{equation}
where $m_i$ and $\phi_i$ are the mass and azimuthal angle of each particle,
respectively. We computed $A_2$ and $A_0$ in cylindrical bins of width
$0.5\,\textrm{kpc}$ from radii of $0$ to $30\,\textrm{kpc}$. We defined the
angle of the bar $\phi_b$ to be twice the angle of the complex number $A_2$ as
measured in the bin extending from a radius of $2.5$ to $3\,\textrm{kpc}$.
After correcting for the periodicity of $\phi_b$, we measured the pattern
speed as one-half the two-sided finite gradient of $\phi_b$ as a function of
time.

@@

\begin{figure*}
    \centering
    \includegraphics[width=18cm]{fig/halo_wake.pdf}
    \caption{The wake excited in the dark matter halo. The dark matter halo wake
    is shown in the \Nbody{} case (\textit{middle panel}) and SMUGGLE case
    (\textit{right panel}) after $2.6\,\textrm{Gyr}$ of evolution. The
    \textit{middle} and \textit{right panels} show a surface density projection
    in the $x$-$y$ plane of the dark matter halo after an axisymmetric average
    has been subtracted. The solid line indicates the direction of the bar while
    the dashed line indicates the direction of the halo wake (both measured by
    taking the second Fourier component within a sphere of all material within a
    radius of $4\,\textrm{kpc}$). The \textit{left panel} shows the time
    evolution of the angle difference between the bar and the halo wake, as
    measured from the second Fourier component. After the first Gyr, the angle
    difference in the SMUGGLE case is smaller than in the \Nbody{} case by about
    a factor of two, reflecting how the dark matter halo in the SMUGGLE case is
    unable to exert as negative a torque on the bar as in the \Nbody{} case.}
    \label{fig:wake}
\end{figure*}

@@

@@

@@

\section{Bar Decomposition}
\label{app:bardecomp}
\begin{figure*}
    \centering
    \includegraphics[width=18cm]{fig/bar_decomp.pdf}
    \caption{Disk decomposition into the barred and unbarred disk. This
    procedure is based on \citet{2016MNRAS.463.1952P}. The \textit{left panel}
    shows a face-on surface density projection through the stellar component of
    the \SMUGGLE{} simulation (disk and bulge) at $t=1\,\textrm{Gyr}$. The
    \textit{middle panel} shows the component of the disk identified as being
    trapped in the bar while the \textit{right panel} shows the component of the
    disk identified as not being trapped in the bar. The fact that the untrapped
    stars form a roughly axisymmetric structure indicates our bar decomposition
    is sufficiently accurate.}
    \label{fig:decomp}
\end{figure*}

@@

@@

>>>>

\newcommand{\RCR}{\ensuremath{R_{\textrm{CR}}}}
\newcommand{\Rot}{\ensuremath{\mathcal{R}}}
\newcommand{\Vc}{\ensuremath{V_{\textrm{c}} }}
\newcommand{\PS}{\ensuremath{\Omega_{\textrm{p}}}}
\newcommand{\Rb}{\ensuremath{R_{\textrm{b}}}}
\newcommand{\kms}{\ensuremath{\textrm{km}/\textrm{s}}}

@@

\section{Methods}
\label{sec:methods}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{fig/ps_torque.pdf}
    \caption{The \textit{upper panel} shows the evolution of the pattern speed.
    As expected, the bar in the \Nbody{} run slows down due to interactions
    between the bar and the dark matter halo. However, the bar in the SMUGGLE
    run does not slow down and instead remains at a constant pattern speed. The
    \textit{lower panel} shows the torque on the bar by different components.
    The solid lines indicate the torque exerted by the halo in both the \Nbody{}
    and SMUGGLE cases. The dashed line is the torque exerted by the gas phase in
    the SMUGGLE run (there is no gas in the \Nbody{} run). Details on how these
    properties are calculated is given in Section~\ref{ssec:bar_analysis}.}\label{fig:prop}
\end{figure}

@@

\subsection{Bar Analysis}
\label{ssec:bar_analysis}
The analysis of various bar properties is performed as follows. First, the
pattern speed is measured from the angle of the second Fourier component. We
measured the second Fourier component by computing,
\begin{equation}
\begin{split}
A_2 &= \sum_i m_i e^{i 2 \phi_i} \\
A_0 &= \sum_i m_i \textrm{,}
\end{split}
\end{equation}
where $m_i$ and $\phi_i$ are the mass and azimuthal angle of each particle,
respectively. We computed $A_2$ and $A_0$ in cylindrical bins of width
$0.5\,\textrm{kpc}$ from radii of $0$ to $30\,\textrm{kpc}$. We defined the
angle of the bar $\phi_b$ to be twice the angle of the complex number $A_2$ as
measured in the bin extending from a radius of $2.5$ to $3\,\textrm{kpc}$.
After correcting for the periodicity of $\phi_b$, we measured the pattern
speed as one-half the two-sided finite gradient of $\phi_b$ as a function of
time.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fig/halo_wake.pdf}
    \caption{The wake excited in the dark matter halo. The dark matter halo wake
    is shown in the \Nbody{} case (\textit{middle panel}) and SMUGGLE case
    (\textit{right panel}) after $2.6\,\textrm{Gyr}$ of evolution. The
    \textit{middle} and \textit{right panels} show a surface density projection
    in the $x$-$y$ plane of the dark matter halo after an axisymmetric average
    has been subtracted. The solid line indicates the direction of the bar while
    the dashed line indicates the direction of the halo wake (both measured by
    taking the second Fourier component within a sphere of all material within a
    radius of $4\,\textrm{kpc}$). The \textit{left panel} shows the time
    evolution of the angle difference between the bar and the halo wake, as
    measured from the second Fourier component. After the first Gyr, the angle
    difference in the SMUGGLE case is smaller than in the \Nbody{} case by about
    a factor of two, reflecting how the dark matter halo in the SMUGGLE case is
    unable to exert as negative a torque on the bar as in the \Nbody{} case.}
    \label{fig:wake}
\end{figure*}

@@

\subsection{Semi-Analytic Model}
We also developed a simple semi-analytic model of a bar-disk-halo system. This
exercise demonstrates that our proposed mechanism follows from a few simplistic
assumptions. Our method follows closely the one developed in
\citet{2022MNRAS.513..768C}. We model the bar-disk-halo system with three
components: a dark matter Hernquist halo \citep{1990ApJ...356..359H}, a
Miyamoto-Nagai disk \citep{1975PASJ...27..533M}, and a pure quadrupole bar
described in \citet{2022MNRAS.513..768C}. The bar and disk components are just
described by their potential, but we integrate the trajectories of test
particles drawn from a Hernquist halo. Note that we do not include the
interactions between these test particles, and so this model is not
self-consistent. We give our chosen parameter values in Appendix~\ref{app:sam}.

We allow the bar in this model to rotate as a solid body. However, we crucially
allow the pattern speed of the bar to freely change with time in accordance with
the torque exerted on the dark matter halo by the bar. In particular, we
subtract the $z$-component of this torque divided by the moment of inertia of
the bar from the pattern speed. Since the radius of corotation \RCR{} is a
parameter in the bar model from \citet{2022MNRAS.513..768C}, we allow the moment
of inertia of the bar to vary with ${\RCR}^2$. To be more precise, we allow
\begin{equation}
I = \frac{I_6}{10^{10} \Msun (\kms)^2} \left( \frac{\RCR}{6\,\textrm{kpc}} \right)^2\textrm{,}
\end{equation}
where $I_6$ is a free parameter chosen by the user. We found that allowing $I_6=8$ is a good approximation to our bar model.

In addition to $I_6$, we allowed for another free parameter - the torque from
the gas phase on the bar, $\tau_{\textrm{gas}}$. This torque is applied to the
bar in the same way as the torque from the halo is applied. The torque is given
in code units ($10^{10}\Msun \left(\kms\right)^2$). We allow this torque to
mildly vary with pattern speed according to the formula,
\begin{equation}
\tau = \tau_{\textrm{gas}} + (40\,\kms/\textrm{kpc} - \Omega_p)\textrm{.}
\end{equation}
We are motivated to do this because of our result in Appendix~\ref{app:varyps}
which shows that more slowly rotating bars have higher gas infall rates and thus
larger positive torques. In the $\tau_{\textrm{gas}}=0$ case, we do not allow
the gas torque to vary with pattern speed.

We show the effect of varying the gas torque $\tau_{\textrm{gas}}$ from $0$ to
$20$ in increments of $2$ in Fig.~\ref{fig:sam}. The solid lines indicate the
semi-analytic model, while the two dashed lines correspond to our fiducial
simulations introduced earlier. For reference, the average torque exerted by the
gas phase on the bar in our fiducial simulation was $11.7$ in code units. We see in Fig.~\ref{fig:sam} that we can reproduce the stability of our fiducial
gas disk simply by including a positive torque on the order of $6$. Our
semi-analytic model with no gas torque can reproduce the pattern speed evolution
of the $N$-body case.

We next take the $\tau_{\textrm{gas}}=0$ and $20$ cases from Fig.~\ref{fig:sam}
and plot the torque evolution. This result, given in Fig.~\ref{fig:sam-torque},
is comparable to the lower panel of Fig.~\ref{fig:prop}. We find that the
$\tau_{\textrm{gas}}=0$ case compares favorably to the \Nbody{} case described
in previous sections. The bar exerts a steady negative torque in this case (blue
line). When a gas torque is included (orange lines), we find that the halo's
torque becomes much weaker, similar to what we found in the SMUGGLE case. The
gas applies a steady positive torque by construction. The exhaustion process is
reproduced in our simple semi-analytic model.

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{fig/samGvar.pdf}
    \caption{A comparison between the pattern speeds of our fiducial disk
    systems and a semi-analytic model. The solid lines indicate the pattern
    speeds assuming a constant positive torque, varying in increments of $2$
    from $0$ to $20$. The dashed lines indicate the pattern speed evolution from
    our fully self-consistent simulations from earlier. We find excellent
    agreement between our fiducial simulations and our semi-analytic model of a
    bar-disk-halo system. Torques are given in code units ($10^{10}\Msun
    \left(\kms\right)^2$).}
    \label{fig:sam}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{fig/sam_torque.pdf}
    \caption{The torque exerted on the bar by various components in our
    semi-analytic model. Solid lines indicate the torque by the halo while the
    dashed line indicates the torque exerted by the gas phase. We chose two
    models with $\tau_{\textrm{gas}}=0$ and $20$, which most closely resemble
    our $N$-body and SMUGGLE disks, respectively. This figure ought to be
    compared to the lower panel of Fig.~\ref{fig:prop}. Overall, we find good
    qualitative agreement.}
    \label{fig:sam-torque}
\end{figure}

@@

\section{Bar Decomposition}
\label{app:bardecomp}
\begin{figure*}
    \centering
    \includegraphics[width=18cm]{fig/bar_decomp.pdf}
    \caption{Disk decomposition into the barred and unbarred disk. This
    procedure is based on \citet{2016MNRAS.463.1952P}. The \textit{left panel}
    shows a face-on surface density projection through the stellar component of
    the \SMUGGLE{} simulation (disk and bulge) at $t=1\,\textrm{Gyr}$. The
    \textit{middle panel} shows the component of the disk identified as being
    trapped in the bar while the \textit{right panel} shows the component of the
    disk identified as not being trapped in the bar. The fact that the untrapped
    stars form a roughly axisymmetric structure indicates our bar decomposition
    is sufficiently accurate. We have computed that $76\%$ of the second Fourier
    component resides in the stars classified as being trapped in the bar.}
    \label{fig:decomp}
\end{figure*}

@@

We computed the second Fourier component $A_2$ for all particles classified as
barred and unbarred. We found that $76\%$ of the total $m=2$ Fourier component
is in the particles classified as barred (i.e.,
$A_{2,\textrm{bar}}/A_{2,\textrm{tot}}\sim0.76$). This indicates good but not
perfect classification. Some of this is probably coming from the $m=4$ component
being classified as unbarred, given that the unbarred component has a somewhat
boxy shape.

@@

\section{Semi-Analytic Model Parameters}
\label{app:sam}
Our semi-analytic model consisted of a three-component bar-disk-halo system. We
describe here the parameters we chose for these components. The parameters of
the disk and halo were chosen to match closely what we used in our fiducial
simulations. The system can thus be understood as being roughly similar to the
Milky Way, though no careful analysis has been performed to ensure the closest
match possible.

For the dark matter halo, we used a Hernquist potential
\citep{1990ApJ...356..359H} with mass $10^{12}\,\Msun$ and a scale length of
$26.2\,\textrm{kpc}$. For the stellar disk, we used a MiyaMoto-Nagai disk
\citep{1975PASJ...27..533M} with mass $4.8\times10^{10}\,\Msun$, radial scale
length of $2.67\,\textrm{kpc}$, and vertical scale length of
$0.32\,\textrm{kpc}$. For the bar, we used the quadrupole potential described in
\citet{2022MNRAS.513..768C}. We used their fiducial parameter values --
specifically, we set $A=0.02$, $b=0.28$, and $v_c = 235\,\kms$. Our initial
pattern speed is always set to $40\,\kms/\textrm{kpc}$.
